\chapter{Preface}

This report covers the master's thesis for the Cybernetics and Robotics study program at the Norwegian University of Science and Technology (NTNU). The work is a continuation of the independent project work conducted during the previous term, and some parts of this report are taken directly from that work. This is stated explicitly in the concerning sections. Furthermore, a lot of the ideas behind this work are inspired by my supervisor Stavdahl's earlier notes and the discussions we have had on the topic.

The simulator used in this project has been developed by Sanfilippo at NTNU with considerable contribution from skilled students. The code for the method developed in this project is written and integrated to the simulator by me. The figures presented throughout the report are also designed by me, unless stated otherwise.

The unfortunate COVID-19 situation that arose early this year has certainly had an impact on the project flow.  %Not so much for the technical part, as all experiments have been performed via the mentioned simulator, but rather the social part. 
There has been a significant lack of theoretical discussions and brainstorming with fellow students, which for me is something that usually has a very positive effect on my work. Luckily, I have still had digital meetings with my supervisor throughout the entire semester  
%Online meetings have also been limiting for good discussions, but have of course been a great 
and I want to praise him for his continuous engagement and availability.


\medskip
\begin {flushright}
  \textit{Atussa Koushan} \\
  \textsc {28.05.2020} \\
  Trondheim, Norway
\end {flushright}



\makeatletter
\ifthenelse{\pdf@strcmp{\languagename}{english}=0}


\makeatother




